---
title: "export_dataset"
author: "James Howison"
date: "3/29/2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(SPARQL)
library(here)
library(data.world)
knitr::opts_chunk$set(echo = TRUE)

softcite_ds <- "https://data.world/jameshowison/software-citations/"
```

softcite_ds <- "http://192.168.99.100:3030/softcite-dataset/"
docker run -p 3030:3030 -e ADMIN_PASSWORD=pw123 -it stain/jena-fuseki 


Data model is:

article has_many selection
selection has_many reference

article has_many code_application (for coded_no_in_text_mentions)
selection has_many code_application
reference has_many code_application

Export as four tables:

1. articles
2. in_text_selection
3. references
4. code_applications (coder, target_id (article|in_text_mention|reference))

articles(id, literature_sample, doi, title etc.)
coders(id, name)
assignments(id, article_id, coder_id, when_coded, selections_found (true|false))
selections(id, assignment_id, page, quote, in_text_or_reference, reference_in_selection)
code_applications (id, selection_id, coder, code, code_was_present, label, memo)


articles(id, literature_sample, doi, title etc.)

assignments(article, coder, when_coded, selections_found (true|false))

in_text_mentions(selection, assignment_id, page, quote)

references(selection, page, quote, mention_type, from_in_text_selection)

code_applications (selection, coder, code, code_was_present, label, memo)

```{r}
prefixes <- "PREFIX bioj: <http://james.howison.name/ontologies/bio-journal-sample#>
PREFIX bioj-cited: <http://james.howison.name/ontologies/bio-journal-sample-citation#>
PREFIX ca: <http://floss.syr.edu/ontologies/2008/4/contentAnalysis.owl#>
PREFIX citec: <http://james.howison.name/ontologies/software-citation-coding#> 
PREFIX dc: <http://dublincore.org/documents/2012/06/14/dcmi-terms/>
PREFIX doap: <http://usefulinc.com/ns/doap#>
PREFIX owl: <http://www.w3.org/2002/07/owl#>
PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
PREFIX vivo: <http://vivoweb.org/ontology/core#>
PREFIX xml: <http://www.w3.org/XML/1998/namespace>
PREFIX xsd: <http://www.w3.org/2001/XMLSchema#>
"
```

```{r}
articles_string <- "SELECT DISTINCT ?article ?article_set ?coder ?no_selections_found ?codable ?standard_article
WHERE {
  ?article rdf:type ?article_set;
           ca:isTargetOf
        [ rdf:type ca:CodeApplication ;
          ca:appliesCode [ rdf:type citec:coded_no_in_text_mentions ;
            citec:isPresent ?no_selections_found; ];
          ca:hasCoder ?coder ]
  OPTIONAL {
  ?article rdf:type ?article_set;
           ca:isTargetOf
        [ rdf:type ca:CodeApplication ;
          ca:appliesCode [ rdf:type citec:codable ;
            citec:isPresent ?codable; ];
          ca:hasCoder ?coder ]
  }
  OPTIONAL {
  ?article rdf:type ?article_set;
           ca:isTargetOf
        [ rdf:type ca:CodeApplication ;
          ca:appliesCode [ rdf:type citec:standard_type ;
            citec:isPresent ?standard_article; ];
          ca:hasCoder ?coder ]
  }
  FILTER(?article_set != bioj:article)
}
"

articles <- data.world::qry_sparql(paste0(prefixes, articles_string)) %>%
  data.world::query(softcite_ds) %>% 
  as_tibble() %>% 
  mutate_at(vars(article, article_set),
            list(~ str_extract(.,"[#/]([^#/]+)$"))) %>%
  mutate_at(vars(article, article_set), list(~ str_sub(.,2)))  %>% 
  write_csv(here::here("data/csv_dataset/softcite_articles.csv"))
```

```{r}
in_text_mention_q <- "SELECT DISTINCT ?selection ?coder ?article ?quote ?tei_quote ?page ?mention_type ?certainty ?memo
WHERE {
	?article citec:has_in_text_mention ?selection .
	?selection citec:full_quote ?quote ;
	           citec:on_pdf_page ?page ;
	           ca:isTargetOf [ ca:hasCoder ?coder ;
                             ca:appliesCode [ rdf:type citec:mention_type ;
                                              rdfs:label ?mention_type ;
                                              ca:certainty ?certainty ;
                                              ca:memo ?memo ]
                            ]
  OPTIONAL { ?selection citec:tei_full_quote ?tei_quote }
}
"

in_text_mentions <- data.world::qry_sparql(paste0(prefixes, in_text_mention_q)) %>%
  data.world::query(softcite_ds) %>% 
#  SPARQL(url = softcite_ds, query = paste0(prefixes, in_text_mention_q), 
#       format = "csv", extra=list(format="csv"))$results %>% 
  as_tibble() %>% 
  mutate_at(vars(article, selection),
            fun(str_extract(.,"[#/]([^#/]+)$"))) %>%
  mutate_at(vars(article,selection), funs(str_sub(.,2))) %>% 
  write_csv(here::here("data/csv_dataset/softcite_in_text_mentions.csv"))

```

```{r}
reference_q <- "SELECT DISTINCT ?reference ?from_in_text_selection ?article ?quote ?coder ?page ?reference_type
WHERE {
  ?article citec:has_in_text_mention ?from_in_text_selection .
  ?from_in_text_selection citec:has_reference ?reference .
	?reference citec:full_quote ?quote ;
	           citec:on_pdf_page ?page ;
	           ca:isTargetOf [ ca:hasCoder ?coder ;
                             ca:appliesCode [ rdf:type citec:reference_type ;
                                              rdfs:label ?reference_type ]
                            ]
}
"

references <- data.world::qry_sparql(paste0(prefixes, reference_q)) %>%
  data.world::query(softcite_ds) %>%
#  SPARQL(url = softcite_ds, query = paste0(prefixes, reference_q), 
#       format = "csv", extra=list(format="csv"))$results %>% 
  as_tibble() %>% 
  mutate_at(vars(article, reference, from_in_text_selection),
            funs(str_extract(.,"[#/]([^#/]+)$"))) %>%
  mutate_at(vars(article, reference, from_in_text_selection), funs(str_sub(.,2))) %>% 
  write_csv(here::here("data/csv_dataset/softcite_references.csv"))
```

code_applications (selection_id, coder, code, code_was_present, label, memo)

```{r}
all_codes_q <- "
SELECT DISTINCT ?selection ?coder ?code ?was_code_present ?code_label
WHERE {
  ?selection ca:isTargetOf [  ca:appliesCode ?ca ;
                              ca:hasCoder ?coder ] .
  ?ca rdf:type ?code ;
      citec:isPresent ?was_code_present .

  OPTIONAL { ?ca rdfs:label ?code_label }
}"

codes <- data.world::qry_sparql(paste0(prefixes, all_codes_q)) %>%
  data.world::query(softcite_ds) %>%
#  SPARQL(url = softcite_ds, query = paste0(prefixes, all_codes_q), 
#      format = "csv", extra=list(format="csv"))$results %>% 
  as_tibble() %>% 
  mutate_at(vars(selection, code),
            funs(str_extract(.,"[#/]([^#/]+)$"))) %>%
  mutate_at(vars(selection, code), funs(str_sub(.,2))) %>% 
  write_csv(here::here("data/csv_dataset/softcite_codes_applied.csv"))
```


```{r}
articles %>% 
  ggplot(aes(x=article_set)) +
  geom_histogram(stat= "count") + 
  facet_grid(no_selections_found ~ .)
```

```{r}
articles %>% 
  group_by(article_set, article) %>% 
  summarize(coder_count = n_distinct(coder)) %>%
  filter(coder_count > 1) %>% 
  ggplot(aes(coder_count)) +
  geom_histogram(stat="count") +
  facet_grid(article_set ~ ., scales = "free")
```

