---
title: "SciSoft Coding Agreement"
subtitle: James Howison
output: html_document
---

```{r}
library(tidyverse)
library(data.world) # loads saved config see quickstart vignette
library(stringdist)

prefixes <- "
PREFIX bioj: <http://james.howison.name/ontologies/bio-journal-sample#>
PREFIX bioj-cited: <http://james.howison.name/ontologies/bio-journal-sample-citation#>
PREFIX ca: <http://floss.syr.edu/ontologies/2008/4/contentAnalysis.owl#>
PREFIX citec: <http://james.howison.name/ontologies/software-citation-coding#> 
PREFIX dc: <http://dublincore.org/documents/2012/06/14/dcmi-terms/>
PREFIX doap: <http://usefulinc.com/ns/doap#>
PREFIX owl: <http://www.w3.org/2002/07/owl#>
PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
PREFIX vivo: <http://vivoweb.org/ontology/core#>
PREFIX xml: <http://www.w3.org/XML/1998/namespace>
PREFIX xsd: <http://www.w3.org/2001/XMLSchema#>
"

softcite_ds = "https://data.world/jameshowison/software-citations/"

# should pull from coding scheme
valid_codes = c("has_supplement",
"has_in_text_mention",
"coded_no_in_text_mentions",
"memo",
"full_quote",
"on_pdf_page",
"spans_pages",
"mention_type",
"software_was_used",
"software_name",
"version_number",
"version_date",
"url",
"creator",
"has_reference",
"reference_type")
```


This shows basic agreement stats for the scisoft content analysis.

```{r}
top_code_query <- data.world::qry_sparql(paste(prefixes,
      "SELECT ?article ?coder ?selection ?full_quote ?on_pdf_page ?spans_pages
WHERE {
    ?article citec:has_in_text_mention ?selection .
    ?selection ca:isTargetOf
        [ rdf:type ca:CodeApplication ;
          ca:hasCoder ?coder ;
          ca:appliesCode [ rdf:type citec:mention_type ]
        ] .
    ?selection citec:full_quote ?full_quote ;
               citec:on_pdf_page ?on_pdf_page ;
               citec:spans_pages ?spans_pages
    }"
))
top_codes <- data.world::query(top_code_query, softcite_ds)
top_codes <- as.tibble(top_codes)
```

SPARQL queries return everything as URLs, but we want just the localPart at the end.

```{r}
top_codes <- top_codes %>%
       mutate_at(vars(article, selection), funs(str_extract(.,"[#/]([^#/]+)$"))) %>%
       mutate_at(vars(article,selection), funs(str_sub(.,2)))
```



```{r}
# Focus on PMC dataset.
top_codes <- top_codes %>% filter(str_detect(article, "PMC"))
```


```{r}
num_articles <- top_codes %>% summarize(article_total = n_distinct(article))
num_articles$article_total

# number of mentions per article
top_codes %>% 
  group_by(article, coder) %>% 
  summarize(selection_count = n_distinct(selection))

```

```{r}
require(RecordLinkage)

top_codes <- top_codes %>% filter(str_detect(article,"PMC"))

links <- compare.linkage(top_codes, top_codes, blockfld = c("article","on_pdf_page"), strcmp = T, exclude = c("selection", "coder", "spans_pages"))
pairs <- as.tibble(links$pairs) %>% 
  select(id1, id2, full_quote) %>% 
  filter(id1 != id2) %>% # exclude self matches. 
  rename(selection_id = id1, match_id = id2, full_quote_match_score = full_quote)

# sort group by full_quote, take top 1, can produce ties
top_matches <- pairs %>% 
  group_by(selection_id) %>% 
  top_n(1, full_quote_match_score) %>% 
  ungroup()
  
# add rowname as integer to top_codes for joining.
sel_lookup <- rownames_to_column(top_codes) 
sel_lookup$id <- as.integer(sel_lookup$rowname)
sel_lookup <- sel_lookup %>% select(id, selection)

# change ids to selection names
top_matches <- left_join(top_matches, sel_lookup, 
                         by = c("selection_id" = "id")) %>% 
  select(-selection_id, selection, match_id, full_quote_match_score)

# now change id in top_match to selection name
top_matches <- left_join(top_matches, sel_lookup, 
                         by = c("match_id" = "id")) %>% 
  rename(selection = selection.x, top_match = selection.y) %>% 
  select(-match_id)

```

Ok, so now avoid matching with own selections.

```{r}
# Function to generate lookups for tibbles
my.lookup <- function(data) {
  data_lookup <- rownames_to_column(data) 
  data_lookup$id <- as.integer(data_lookup$rowname)
  data_lookup <- data_lookup %>% select(id, selection)
  return(data_lookup)
}

# Function to return matching results from compare.linkage in friendly way
my.compare.linkage <- function(data1, data2, ...) {
  linkage <- compare.linkage(data1, data2, ...)
  pairs <- linkage$pairs %>% 
    select(-is_match, -on_pdf_page) %>% 
    rename(match_score = full_quote)
  
  # swap ids.
  data1_lookup <- my.lookup(data1)
  data2_lookup <- my.lookup(data2)
  
  pairs <- pairs %>% 
    left_join(data2_lookup, by = c("id2" = "id")) %>% 
    rename(poss_match = selection) %>% 
    left_join(data1_lookup, by = c("id1" = "id")) %>% 
    select(-id1, -id2)

}

my.stringdist <- function(string1, string2) {
  return(stringsim(string1, string2, method = "jaccard"))
}
```


```{r}
# remove articles with only 1 coder.
# This helps to avoid NAs in the possible matches.
multi_coders <- top_codes %>% 
  group_by(article) %>% 
  mutate(num_coder = n_distinct(coder)) %>% 
  ungroup() %>%
  filter(num_coder > 1)

get_possible_matches <- function(curr_article, curr_coder, data) {
  m <- data %>% filter(article == curr_article, coder != curr_coder )
  select(m, -article, -coder, -num_coder)
}


poss_matches <- multi_coders %>% 
  group_by(article, coder) %>%
  summarize(poss_matches = list(get_possible_matches(article, coder, multi_coders))) 

data_for_matching <- multi_coders %>% 
  select(-num_coder) %>% 
  group_by(article, coder) %>% 
  nest(.key = "selections") 


# removes any NAs
grouped_data <- left_join(poss_matches, data_for_matching)

results <- grouped_data %>% 
  mutate(pairs = map2(selections, poss_matches,
                      my.compare.linkage, 
                      strcmp=T, exclude=c("selection", "spans_pages"),
                      strcmpfun = my.stringdist # method defined above
                      )
        )

unnested_results <- results %>% 
  select(-selections, -poss_matches) %>% 
  unnest() %>% 
  select(article, coder, selection, poss_match, match_score) %>%
  ungroup()
```

Now we can rejoin with top codes, or get highest match for each, or a threshold. Some don't (or won't after threshold) have any matches. Should produce an NA when joining with top_codes.

```{r}

# How many have a high match?
filter(unnested_results, match_score > 0.8) %>% tally()

# Which have more than one high match? Likely these are multiples mentions in a sentence.
filter(unnested_results, match_score == 1) %>% group_by(selection) %>% tally() %>% filter(n > 1) %>% ungroup() %>% tally()

# Let's see the full quotes of best matches, have to join on two column
best_matches <- unnested_results %>% group_by(selection) %>% top_n(n=1, match_score) %>% 
  left_join(select(top_codes, selection, full_quote), by = "selection") %>% 
  left_join(select(top_codes, selection, full_quote), by = c("poss_match" = "selection"))

# Put most likely match, over threshold, into original frame.
# This pipeline brings the best match them into top_codes.
top_codes_with_matches <- unnested_results %>% 
  filter(match_score > .8) %>% 
  group_by(selection) %>% 
  top_n(n=1, match_score) %>% 
  select(selection, poss_match) %>% 
  right_join(top_codes, by = "selection") %>% 
  select(article, coder, selection, everything())

# put full_quote from top_code into the unnested_results table, for ease of visual comparison.
# selection and poss_match
unnested_results_with_quotes <- unnested_results %>% 
  left_join(select(top_codes, selection, full_quote), by = "selection") %>%
  rename(full_quote_selection = full_quote) %>% 
  left_join(select(top_codes, selection, full_quote), by = c("poss_match" = "selection")) %>%
  rename(full_quote_poss_match = full_quote)
```


What we actually want to know is: for each article that had at least 2 coders, what was the percentage agreement in identifying selections?

percentage agreement = count of matched / all selections.

```{r}
# selections from articles with more than 1 coder.
multi_coder_selections <- top_codes_with_matches %>% 
  group_by(article) %>% 
  mutate(num_coder = n_distinct(coder)) %>% 
  ungroup() %>%
  filter(num_coder > 1) 

# total number of selections by either coder, matched and un-matched.
unmatched_counts <- multi_coder_selections %>% 
  group_by(article) %>% 
  filter(is.na(poss_match)) %>% 
  summarize(unmatched_count = n_distinct(selection))

# total number of matched selections (A -> B, B -> A counts as 2.)
matched_counts <- multi_coder_selections %>% 
  filter(!is.na(poss_match)) %>% 
  group_by(article) %>% 
  summarize(matched_count = n_distinct(selection))

percent_agree <- multi_coder_selections %>% distinct(article)
#percent_agree$matched_count <- 0
#percent_agree$unmatched_count <- 0

percent_agree <- percent_agree %>% 
  left_join(unmatched_counts, by = "article") %>%
  left_join(matched_counts, by = "article") %>%
  mutate(unmatched_count = replace(unmatched_count, is.na(unmatched_count), 0),
         matched_count = replace(matched_count, is.na(matched_count), 0)
        ) %>% 
  mutate(percent_agree = matched_count / ( matched_count + unmatched_count )) %>% 
  filter(percent_agree < 1)


ggplot(percent_agree, aes(x=reorder(article, percent_agree), y=percent_agree)) + geom_bar(stat="identity") + scale_y_continuous(labels=scales::percent) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# lowest agreement
arrange(percent_agree, percent_agree)
```
```{r}
data <- unnested_results_with_quotes %>% filter(article == "PMC2816832") %>% distinct(selection, poss_match, .keep_all=T) 
View(data %>% arrange(match_score))
```


```{r}
lower_code_query <- data.world::qry_sparql(paste(prefixes,
      "SELECT ?article ?coder ?selection ?code ?present ?label
WHERE {
    ?article citec:has_in_text_mention ?selection .
    ?selection ca:isTargetOf ?ca .
    ?ca ca:hasCoder ?coder ;
        ca:appliesCode ?code_block .
    ?code_block rdf:type ?code ;
                citec:isPresent ?present .
OPTIONAL { ?code_block rdfs:label ?label }
}"
))
lower_codes <- data.world::query(lower_code_query, softcite_ds)
lower_codes <- as.tibble(lower_codes)
lower_codes <- lower_codes %>% mutate(present = (present == "true"))
```

Join them together, vertically. Convert present column to boolean

```{r}
df <- rbind(lower_codes, top_codes)
df <- distinct(df)
```

```{r}
df$code <- readr::parse_factor(df$code, levels = valid_codes)
```

Now obtain a graph showing counts of codes per user for each article
```{r, echo=FALSE}
p <- df %>% 
  group_by(article, coder) %>% 
 # summarize(num_mentions = n()) %>% 
  ggplot(aes(x = coder)) +
  geom_bar() +
  scale_y_continuous(breaks = c(2,4,6,8), minor_breaks = FALSE) +
  facet_wrap_paginate(~ article, ncol = 2, nrow = 2, page = 2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
library(Biostrings)
#pairwiseAlignment(pattern = c("succeed", "precede"), subject = "supersede")

aligns <- pairwiseAlignment(pattern = c("The GenomeRNAi database", "The GenomeRNAi database provides access"), subject = "The GenomeRNAi database provides access topublish", type = "overlap")

summary(aligns)

compareStrings(pattern = c("The GenomeRNAi database"), subject = "The GenomeRNAi database provides access topublish")

nmatch(c("The GenomeRNAi database","The GenomeRNAi database provides access topublish"))
```


```{r}

quote1 = "the cat sat on the mat"
quote1list <- quote1 %>% str_to_lower() %>% str_split(" ") %>% unlist()
quote2 <- "on the mat"
quote2list <- quote2 %>% str_to_lower() %>% str_split(" ") %>% unlist()

quote1list
quote2list
overlap <- intersect(quote1list, quote2list)
diffa <- setdiff(quote1list, quote2list)
diffb <- setdiff(quote2list, quote1list)

length(overlap) / (length(overlap) + length(c(diffa, diffb)))
length(overlap) / 


selection_overlap <- function(quote1, quote2) {
  # normalize and split on spaces
  quote1list <- quote1 %>% str_to_lower() %>% str_split(" ") %>% unlist()
  quote2list <- quote2 %>% str_to_lower() %>% str_split(" ") %>% unlist()
  overlap <- intersect(quote1list, quote2list)
  return(length(overlap) / length(union(quote1list, quote2list)))
}

selection_overlap("the cat sat on the mat", "on the mat")
selection_overlap("the cat sat on the mat", "doctor who")
selection_overlap("the cat sat on the mat", "because two the cat sat on")


```

