---
title: "Softcite Processing Summary"
author: "Fan Du"
date: "9/11/2020"
output: html_document
---

This document is built in comparison with:
https://github.com/ourresearch/software-mentions/blob/master/doc/corpus-counts.md

Difference may come from differed rules used for calculation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(here)
library(xml2)
library(janitor)
library(knitr)
library(kableExtra)

# load csv data
csv_articles <- read_csv(here("data/csv_dataset/softcite_articles.csv"))
csv_mentions <- read_csv(here("data/csv_dataset/softcite_in_text_mentions.csv"))
csv_codes <- read_csv(here("data/csv_dataset/softcite_codes_applied.csv"))
```

**Number of articles annotated in the CSV data:**
```{r, echo=TRUE}
csv_articles %>% 
  filter(article_set %in% c("pmc_article", "econ_article")) %>% 
  group_by(article) %>% 
  # 5,553 articles annotated
  #filter(codable == TRUE | is.na(codable)) %>% 
  # 5,549 after excluding non-codable articles
  #filter(standard_article == TRUE | is.na(standard_article)) %>% 
  # 4,645 after excluding non-standard articles
  tabyl(no_selections_found) %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting() %>% 
  kable() %>% kable_styling()
```

```{r, echo=TRUE, results='hide'}
raw_software_annotations <- csv_articles %>% 
  filter(article_set %in% c("pmc_article", "econ_article")) %>% 
  filter(no_selections_found == F) %>% 
  left_join(csv_mentions, by=c("article", "coder")) %>% 
  filter(mention_type == "software") %>% 
  select(-codable, -standard_article, -quote, -tei_quote, -memo, -certainty, -page) %>% 
  # TODO: there might be duplicated values in the fields dropped
  # make sure that every record is uniquely identified
  distinct()
  # 6,206 annotations of software entities (i.e. unique selections)
  # NOTE: some selections may not be uniquely identified during annotation but very rare. Estimated difference within ~10

raw_software_names <- raw_software_annotations %>% 
  left_join(csv_codes, by=c("selection", "coder")) %>% 
  filter(code == "software_name") 
  # 6,208 named and non-named software annotations

raw_software_attributes <- raw_software_annotations %>% 
  left_join(csv_codes, by=c("selection", "coder")) %>% 
  filter(code %in% c("software_was_used", "creator", "version_date", "version_number", "url") & was_code_present == T) 
  # 9,510 annotations of software attributes
  # 6208 + 9510 = 15718 labels in total

raw_annotation_pairs <- raw_software_names %>% 
  left_join(raw_software_attributes, by=c("selection", "coder")) %>% 
  select(-code.x, -article.y, -article_set.y, -no_selections_found.y, -mention_type.y) %>% 
  rename(article=article.x, article_set=article_set.x, no_selections_found=no_selections_found.x, mention_type=mention_type.x, named=was_code_present.x, software_name=code_label.x, attribute=code.y, was_attr_present=was_code_present.y, attr_detail=code_label.y)

top_annotators <- raw_annotation_pairs %>% 
  group_by(article, coder) %>% 
  summarise(n_anno_per_coder=sum(n_distinct(software_name)+n_distinct(attr_detail))) %>% 
  top_n(n=1, wt=n_anno_per_coder)
# top annotator in 1,618 articles

top_annotators %>% 
  left_join(raw_software_names, by=c("article", "coder")) %>% 
  distinct()
# 5,658 software entities annotated by top annotators

top_annotators %>% 
  left_join(raw_software_attributes, by=c("article", "coder")) %>% 
  distinct()
# 8,768 software attributes annotation made by top annotators
# 5658 + 8768 = 14426 lables in total produced by top annotators in each annotated article
# TODO: check missing values in each field.

# The definition of the largest run has not been clarified yet. Here I use the annotator at one particular article who produced the most annotations of software entity and attributes.
```

```{r, echo=TRUE, results='hide'}

tei_doc <- read_xml("https://raw.githubusercontent.com/ourresearch/software-mentions/master/resources/dataset/software/corpus/softcite_corpus-full.tei.xml") %>% xml_ns_strip()

xml_find_num(tei_doc, xpath="count(//fileDesc[@xml:id])")
# 4,971 articles in softcite_corpus-full.tei.xml

xml_find_num(tei_doc, xpath="count(//body[not(node())])")
# 3,544 articles with zero mentions
xml_find_num(tei_doc, xpath="count(//body[.//text()])")
# 1,427 articles with mentions

xml_find_num(tei_doc, xpath="count(//rs[@type='software'])")
# 5,171 annotations of software entity
xml_find_num(tei_doc, xpath="count(//rs)")
# 8,335 annotations in total, all are reviewed by curators
# which means 8335-5171=3164 annotations of software attributes

xml_find_num(tei_doc, xpath="count(//rs[@type='software' and @resp='#curator'])")
# 741 annotations of software entity edited by curator
xml_find_num(tei_doc, xpath="count(//rs[@resp='#curator'])")
# 2,043 annotations edited by curator in total
```

**Process summary statistics break down by steps (as of May 30, 2020):**
(Estimated difference < ~10)

|steps|software_annotation_count|annotation_total_count|articles_with_annotations|articles_with_zero_annotations|article_total_count|
|-----|----------------------|----------------------|-------------------------|------------------------------|-------------------|
|Raw annotation runs|6208|15718|1823|3730|5553|
|Raw annotation largest run|5658|14426|1823|3730|5553|
|Annotations located in PDFs|-|-|-|-|-|
|Annotations reviewed by curator|5170|8335|1427|3544|4971|
|Annotations edited by curator|741|2043|1427|3544|4971|


Notes:

1. Statistics in the first 2 rows are calculated from CSV data. 
It should be noted that the definition of the largest run has not been clarified yet. Here I use the annotator at one particular article who produced the most annotations of software entity and attributes.

2. Statistics in the 4th and 5th row are queried from the current version of softcite TEI corpus.

